---
title: "Lab 2: Linear Regression"
author: "Your Name"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

Before beginning this lab, you may want to read or skim Section 3.6 in the textbook, or watch [this video*](https://www.youtube.com/watch?v=W0WILgmAGao&list=PL5-da3qGB5IBSSCPANhTgrw82ws7w_or9&index=6). That section includes lots of helpful commands to complete the exercises in this lab. You may also like to check out the file `linear_regression.Rmd` available in the Week 2 module on Canvas.

*Note that the authors use a regular "R Script" rather than a RMarkdown file (which is what this file is). Also, many of their graphics are produced with "base R" rather than the prettier `ggplot` commands we learned in the first lab.

---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(ISLR2)
library(broom)
```

This lab will use the `Auto` data set, which is included in the ISLR2 R-package. Use the command `names()` to find the variables of the data set. Try the command `?Auto` to see a full description of the data set and every variable -- this command must be used in the console.

```{r}
names(Auto)
```

## Simple Linear Regression
<!-- Exercise 3.8, page 123 -->

### Exercise 1

Use the `lm()` function to perform a simple linear regression with `mpg` as the response and `horsepower` as the predictor. Use the `summary()` and `tidy()` functions to print the results.

For the `tidy()` function, you'll need to load the `broom` package -- best practice is to add it to the other `library()` commands at the top of your RMarkdown document.

```{r}
ex1 <- lm(mpg ~ horsepower, data = Auto)
summary(ex1)
print(tidy(ex1))
<<<<<<< Updated upstream
=======

>>>>>>> Stashed changes
```

Use your model to answer the following questions:

a. Is there a statistically significant relationship between the predictor and the response?
b. How strong is the relationship between the predictor and the response?
c. Is the relationship between the predictor and the response positive or negative?
d. What is the predicted `mpg` associated with a `horsepower` of 98? What are the associated 95% confidence and prediction intervals?

```{r}

```

### Exercise 2 
Make two plots to visualize your regression model:

a. Plot the response and the predictor using base R. Use the `abline()` function to display the least squares regression line.
b. Use `ggplot` and the `geom_smooth` layer.

```{r}

```


### Exercise 3
You will now evaluate the least squares regression fit of your model. For each of the following "Potential Problems", compute a measurement statistic, produce a plot, or both. Then comment on whether this might be an issue for your model from above. (You can review these on page 92 of our textbook.)

a. Non-linearity
b. Correlation of error terms
c. Non-constant variance of error terms
d. Not Normal Residuals
e. Outliers
f. High-leverage points
g. Collinearity 


## Multiple Linear Regression
<!-- Exercise 3.9, page 123-4 -->

### Exercise 4
Produce a scatterplot matrix which includes all of the variables in the data set. Also, compute the matrix of correlations between all of the variables.

```{r}

```

Now try the command `ggpairs()` instead.

```{r}

```


### Exercise 5
Use the `lm()` function to perform a multiple linear regression with `mpg` as the response and all other variables except `name` as the predictors. Print the results and use them to answer the following:

a. Is there a statistically significant relationship between the predictors and the response as a whole?
b. Which predictors appear to have a statistically significant relationship to the response in this model?
c. What does the coefficient for the `year` variable suggest?

```{r}

```

### Exercise 6
There is a problem with your multiple linear regression model. Examine the `origin` variable and it's description (use `?Auto` in the console) to diagnose the problem.

Replace occurences of "origin#" in the code below to create a new variable that accurately encodes the categorical nature of `origin`.

```{r}
Auto <- Auto %>%
  mutate(origin_cat = as.factor(
    ifelse( origin == 1 | origin == 2,  
      ifelse( origin == 1, "origin1", "origin2" )
    , "origin3" )
    )
  )
```

Explore your new variable with the `contrasts()` function, replace your model from Exercise 7 with a new multiple regression model that incorporates your new variable instead of `origin` as a predictor.


### Exercise 7
Repeat the steps in Exercise 3, above to evaluate the least squares regression fit of your model with respect to each of the following "Potential Problems":

a. Non-linearity
b. Correlation of error terms
c. Non-constant variance of error terms
d. Not Normal Residuals
e. Outliers
f. High-leverage points
g. Collinearity 

```{r}

```

### Exercise 8
Use the `*` or `:` symbols to fit linear regression models with interaction effects. Do any interactions appear to be statistically significant?

```{r}

```

### Exercise 9
Try a few different transformations of variable such as $\log(X), \sqrt{X}, X^2$. Comment on your findings.

```{r}

```

### Exercise 10
Write a function that takes a data frame as input, runs a linear regression, and outputs diagnostic plots to help in evaluating the "Potential Problems" of Exercises 3 and 7.


