---
title: "Lab0: Intro to R"
author: "Xander"
date: "`r Sys.Date()`"
output: openintro::lab_report
format:
  html:
    code-fold: true
    code-summary: "Show the code"
---

Before beginning this lab, you may want to read or skip Section 2.3 in the textbook. That section includes lots of helpful commands to complete the exercises in this lab.

---

```{R libraries, message=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(openintro)
library(ISLR2)
```

This exercise relates to the College dataset, which is included in the ISLR2 R-package. It contains a number of variables for 777 different universities and colleges in the US. The variables are

* Private : Public/private indicator
* Apps : Number of applications received
* Accept : Number of applications accepted
* Enroll : Number of new students enrolled
* Top10perc : New students from top 10% of high school class
* Top25perc : New students from top 25% of high school class
* F.Undergrad : Number of full-time undergraduates
* P.Undergrad : Number of part-time undergraduates
* Outstate : Out-of-state tuition
* Room.Board : Room and board costs
* Books : Estimated book costs
* Personal : Estimated personal spending
* PhD : Percent of faculty with Ph.D.'s
* Terminal : Percent of faculty with terminal degree
* S.F.Ratio : Student/Faculty ratio
* perc.alumni : Percent of alumni who donate
* Expend : Instructional expenditure per student
* Grad.Rate : Graduation rate

### Exercise 1

First, use the `glimpse()` and `head()` commands to see the variables and some sample values. How are these commands different? How are they the same?

```{r}
glimpse(College)
head(College)
```

Next, use the `summary()` function to produce a numerical summary of the variables in the data set.

```{r}
summary(College)
```

### Exercise 2

Use the `pairs()` function to produce a *scatterplot matrix* of the first ten columns or variables of the data. Recall that you can reference the first ten columns of a matrix A using `A[,1:10]`

```{r, fig.width=12, fig.height=12}
pairs(College[, 1:10])
```

### Exercise 3

Produce side-by-side boxplots showing the distribution of numbers of students from out of state (`Outstate`) in both public and private institutions  (`Private`).

There are two ways to complete this exercise:

* You could use the `plot()` command as described on page 50.
* You could use the `ggplot` package (part of `tidyverse`) and the `+ geom_boxplot()` layer. For more on `ggplot` see the [visualization cheat sheet](https://canvas.hamline.edu/courses/15410/modules/items/689958) on Canvas.

```{r, fig.width=10, fig.height=8}
ex3 <- ggplot(College, aes(x = Private, y = Outstate)) +
  geom_boxplot() +
  ggtitle("ex3")

print(ex3)
```

### Exercise 4

Create a new qualitative variable, call it `Elite`, by binning the `Top10perc` variable. We are going to divide universities into two groups based on whether or not the proportion of students coming from the top 10% of their high school classes exceeds 50%.

The command shown here uses the syntax of the `tidyverse` package.

```{r}
elite_college <- College %>%
  mutate(Elite = as_factor(ifelse(Top10perc > 50, "Yes", "No")))

head(elite_college)
```

Use the `summary()` function to see how many `Elite` universities there are.

```{r}
summary(elite_college)
```

Produce side-by-side boxplots of `Outstate` versus `Elite`. See Exercise 3 for suggestions on how to proceed.

```{r, fig.width=10, fig.height=8}
ex4 <- ggplot(elite_college, aes(x = Private, y = Outstate)) +
  geom_boxplot() +
  ggtitle("ex3")

print(ex3)
```

### Exercise 5

Choose at least two quantitative variables in the data set. For each one, produce at least three histograms with differing number of bins. (In total, you should have at least 6 histograms.)

There are two ways to complete this exercise:

* You could use the `hist()` command as described on page 51.
* You could use the `ggplot` package (part of `tidyverse`) and the `+ geom_histogram()` layer. For more on `ggplot` see the [visualization cheat sheet](https://canvas.hamline.edu/courses/15410/modules/items/689958) on Canvas.

```{r, fig.width=10, fig.height=8}
outstate_hist_10 <- ggplot(College, aes(x = Outstate)) +
  geom_histogram(bins = 10, fill = "#ff7eb6", color = "#000000") +
  ggtitle("Outstate Tuition - 10 Bins") +
  ggtitle("ex5")

outstate_hist_20 <- ggplot(College, aes(x = Outstate)) +
  geom_histogram(bins = 20, fill = "#82cfff", color = "#000000") +
  ggtitle("Outstate Tuition - 20 Bins") +
  ggtitle("ex5")

outstate_hist_30 <- ggplot(College, aes(x = Outstate)) +
  geom_histogram(bins = 30, fill = "#42be65", color = "#000000") +
  ggtitle("Outstate Tuition - 30 Bins") +
  ggtitle("ex5")

print(outstate_hist_10)
print(outstate_hist_20)
print(outstate_hist_30)

room_board_hist_10 <- ggplot(College, aes(x = Room.Board)) +
  geom_histogram(bins = 10, fill = "#ff7eb6", color = "#000000") +
  ggtitle("Room and Board Costs - 10 Bins") +
  ggtitle("ex5")

room_board_hist_20 <- ggplot(College, aes(x = Room.Board)) +
  geom_histogram(bins = 20, fill = "#82cfff", color = "#000000") +
  ggtitle("Room and Board Costs - 20 Bins") +
  ggtitle("ex5")

room_board_hist_30 <- ggplot(College, aes(x = Room.Board)) +
  geom_histogram(bins = 30, fill = "#42be65", color = "#000000") +
  ggtitle("Room and Board Costs - 30 Bins") +
  ggtitle("ex5")

print(room_board_hist_10)
print(room_board_hist_20)
print(room_board_hist_30)
```

### Exercise 6

Using your *scatterplot matrix* from Exercise 2, determine which variable is the strongest predictor for the number of enrolled students (`Enroll`). Create a larger scatterplot showing this relationship and explain why what this means in the context of the data.

There are two ways to complete the plotting in this exercise:

* You could use the `plot()` command as described on page 50.
* You could use the `ggplot` package (part of `tidyverse`) and the `+ geom_point()` layer. For more on `ggplot` see the [visualization cheat sheet](https://canvas.hamline.edu/courses/15410/modules/items/689958) on Canvas.

```{r}
library(ggplot2)
library(GGally)

college_data <- College[, 1:10]

ex6 <- ggplot(college_data, aes_string(x = "Outstate", y = "Top10perc")) +
  geom_point() +
  geom_smooth(method = "lm", color = "#6929c4") +
  labs(x = "Outstate", y = "Top10perc") +
  theme_minimal() +
  ggtitle("ex6")

print(ex6)
```

### Exercise 7

Examine another relationship between two variables that interest you, similar to what you did in Exercise 6.

```{r}
ex7 <- ggplot(College, aes_string(x = "Outstate", y = "Grad.Rate")) +
  geom_point() +
  geom_smooth(method = "lm", color = "#6929c4") +
  labs(x = "Outstate", y = "Grad.Rate") +
  theme_minimal() +
  ggtitle("ex7")

print(ex7)
```

### Continue exploring the data, and provide a brief summary of what you discover

From this graph we can see that schools with a higher out of state tuition also have a higher room and board cost.

```{r}
continue <- ggplot(College, aes_string(x = "Outstate", y = "Room.Board")) +
  geom_point() +
  geom_smooth(method = "lm", color = "#6929c4") +
  labs(x = "Outstate", y = "Room.Board") +
  theme_minimal() +
  ggtitle("continue explore")

print(continue)
```
